 

### 2019-03-03 Manuscript/Manuscript.Rmd
Manuscript Update

In social sciences the use of cylindrical data is very common. Such data consist
of a linear and a circular component. @gurtman2011reasoning refers to such data
as vectors, with a directional measure (i.e., the circular component) and a
measure indicating the magnitude (i.e., linear component). Many established
models in psychology are often referred to as circular or circumplex models, but
those models are cylindrical. Examples of such cylindrical models are the
interpersonal circle/circumplex
[@leary1957;@wiggins1996history;@wubbels2006interpersonal], the circumplex of
affect [@russell1980circumplex], the circumplex of human emotion
[@plutchik1997general] or the model of human values [@schwartz1992values].
\newline
\indent Also, many of the more recent types of data that are studied in psychology are
cylindrical. For example, research on human navigation uses data where distance
(i.e., linear component) and direction (i.e., the linear component) are of
interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
consist of both the direction (i.e., the circular variable) and the duration
(i.e., the linear variable)(e.g., @rayner200935th). Apart from the social
sciences data with a circular and linear outcome more commonly occur in
meteorology [@garcia2013exploring], ecology [@garcia2014test] or
marine research [@lagona2015hidden]
\newline
\indent Up until now researchers studying cylindrical data had to rely on linear
statistical methods to analyze their research results. However, lately more and more
of these researchers acknowledge that linear methods are not sufficient and call
for new methods [@gurtman2011reasoning; @pennings2017complexity;
@wright2009integrating] that take into account both the circular and the linear
component of these data.
\newline
\indent The aim in the present paper is twofold. Firstly, we we intend to fill
the above mentioned gap in the literature by showing the use of cylindrical
models can benefit the analysis of circumplex data and cylindrical data in
psychology in general. More specifically we will show these benefits for
interpersonal teacher data from the field of educational psychology. Apart from
modelling the dependence between the linear and circular component of a
cylindrical variable we would also like to predict the two components from a set
of covariates in a regression model. Our second aim therefore is to adapt
several existing cylindrical models in such a way that they include a regression
of both the linear and circular component of a cylindrical variable onto a set
of covariates. From now on we will therefore refer to the components of the
cylindrical variable as outcome components. These adapted circular models are
then used  to analyse the teacher data.

@@

Data that consist of a variable and a circular variable is called
cylindrical data. A circular variable is different from the linear variable in
the sense that it is measured on a different scale. Figure \ref{circline} shows
the difference between a circular scale (right) and a linear scale (left). The
most important difference is that on a circular scale the datapoints 0\(^\circ\)
and 360\(^\circ\) are connected and in fact represent the same number while on a
linear scale the two ends, \(-\infty\) and \(\infty\), are not connected and
consequently the values 0\(^\circ\) and 360\(^\circ\) are located on different
places on the scale. Both circular data and cylindrical data require special
analysis methods due to this periodicity in the scale of a circular variable
(see e.g. @fisher1995statistical for an introduction to circular data and
@mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
elaborate overview).\newline
\indent In the literature, several methods have been put forward to model the
relation between the linear and circular component of a cylindrical variable.
Some of these are based on regressing the linear component onto the circular
component using the following type of relation: \[y = \beta_0 +
\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
linear component and \(\theta\) the circular component [@mardia1978model;
@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a
different way, e.g. by specifying a multivariate model for several linear and
circular variables and modelling their covariance matrix
[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For
example, @abe2017tractable introduce a cylindrical distribution based on a
Weibull distribution for the linear component and a sine-skewed von Mises
distribution for the circular component and link these through their respective
shape and concentration parameters. However, none of the methods that have been
proposed thus far include additional covariates onto which both the circular and
linear component are regressed.

@@

two-argument arctangent function in \eqref{PredVal}\footnote{The selection of

@@

bivariate space.}, where \(A\) represents a score on the Agency dimension and
\(C\) represents a score on the Communion dimension. Note that when placing a

@@

The teacher data was collected between 2010 and 2015 and contains several
repeated measures on the IPC of 161 teachers. Measurements were obtained using
the QTI and taken in different years and classes. For this paper we only
consider one measurement, the first occasion (2010) and largest class if data
for multiple classes were available. This results in a sample of 151 teachers.
In addition to the type of interpersonal behavior (IPC), the circular outcome,
and the strength of interpersonal behavior (IPC strength), the linear outcome, a
teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
covariate in the analysis. In previous research, in psychology and education it
has been shown that higher self-efficacy is related to the quality of
interpersonal interactions [@locke2007selfefficacy; @want2018selfefficacy].
After listwise deletion of missings ($3$ in total, only for the self-efficacy)
we have a sample of 148 teachers. Table \ref{Tableteacherdescriptives} shows
descriptives for the dataset. For the circular variable IPC we show sample
estimates for the circular mean $\bar{\theta}$ and concentration $\hat{\rho}$.
The circular concentration lies between 0, meaning the data is not concentrated
at all \emph{i.e.} spread over the entire circle, and 1, meaning all data is
concentrated at a single point on the circle The population values of these
parameters are usually, and also in this paper, referred to as $\mu$ (circular
location) and $\kappa$ (circular concentration). For the linear variables
(strength IPC and SE) we show sample estimates of the linear mean and standard
deviation (sd). Figure \ref{dataplot} is a scatterplot showing the relation between
the linear and circular outcome of the teacher data for teachers with low SE
(below 1 sd below the mean), average SE (between 1 sd below and 1 sd above the
mean) and high SE (above 1 sd above the mean).

@@

One of the goals of this paper is to show the benefits of cylindrical methods
for the analysis of circumplex data and cylindrical data in psychology in
general. To do so we decide to focus on four cylindrical models. The models were
selected for their relatively low complexity and the ease with which a
regression structure could be incorporated. But also because they show different
ways of modelling the the linear and circular outcome and thereby illustrate a
wider range of cylindrical models available in the literature. The
cylindrical models contain a set of $q$ predictors $\boldsymbol{x} = x_1, \dots,
x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_q$ for the linear and
circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
based on a construction by @mastrantonio2015bayesian, while the other models are
extensions of the models from @abe2017tractable and @mastrantonio2018joint. The
four cylindrical  models are introduced separately in the subsections below.
However, to provide a more succinct overview and comparison of the four models, Table
\ref{TableModels} gives an overview of the similarities and differences between
the models.

\begin{sidewaystable}[h]
\centering
\caption{Comparison of the four cylindrical regression models} 
\begin{tabular}{lllll}
  \noalign{\smallskip}\hline\noalign{\smallskip}
\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
$\Theta$ & &&&\\
$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
$\:\:$Shape & symmetric, & assymetric, & assymetric, & assymetric, \\
            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
$Y$& &&&\\
$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(0, + \infty)$\\
$\:\:$Shape & symmetric, & symmetric, & assymetric, & assymetric, \\
            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
$\Theta$-$Y$ dependence &                                   &                                   & & \\
                        & $y$ regressed on                  & $y$ regressed on                  & $\alpha$ and $\kappa$ & multivariate \\
                        & $\sin(\theta)$ and $\cos(\theta)$ & $\sin(\theta)$ and $\cos(\theta)$ & & distribution\\\hline
\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\

@@

circular outcome, it is a variable that in contrast to \(\Theta\) is not

@@

\indent For both the PN and GPN distribution the circular location parameter

@@

integretion or Monte Carlo methods because a closed form expression for the mean

@@

\ref{circregline}). By default, the parameters from the PN and GPN model do not

@@

& \xi\\ \xi & 1 \end{bmatrix}\), allowing for multimodality and
assymetry/skewness.\newline

@@

the scores on the Agency an Communion dimensions respectively.\newline

@@

\begin{figure}[]
  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
  \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to. The square indicates the inflection point of the regression live.}
  \label{circregline}
\end{figure}

@@

This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. It concerns a combination
of a Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
\(\alpha\), for the linear outcome and a sine-skewed von Mises distribution,
with location parameter \(\mu\in [0, 2\pi)\), concentration parameter
\(\kappa>0\) and skewness \(\lambda \in [-1,1]\), for the circular outcome. In
contrast to the CL-PN and CL-GPN models, the linear outcome \(Y\) is in this
model defined only on the positive real half-line \([0, + \infty)\) and thus can
not be negative.\newline
\indent In this model we predict the linear scale parameter and circular
location parameter, both of which we can express in terms of
covariates: \(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\gamma_0\) is the intercept. The parameter \(\beta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\).\newline
\indent For the teacher data, the predicted values for the circular outcome in
the Abe-Ley model are: \[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear outcome.
The conditional distribution for the linear outcome is Weibull, meaning that we
can use methods from survival analsis to interpret the effect of a predictor. In
survival analysis a 'survival' function is used in which time is plotted against
the probability of survival of subjects suffering from a specific medical
condition. In the teacher data we can thus compute the probability of a teacher
having a specific strength on the IPC. This probability is computed using the
'survival-function' defined as  \[\exp(-\alpha
y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the survival
function we also see that the circular concentration parameter \(\kappa\) and
the linear shape parameter \(\alpha\) regulate the circular-linear dependence in
the Abe-Ley model. The Abe-Ley model thus allows us to assess the average type
and strength of interpersonal behavior through the parameters $\beta_{0}$ and
$\gamma_0$ as well as the effect of self-efficacy on type and strength of
teacher behavior through the parameters, $\beta_{1}$ and $\gamma_1$.\newline
\indent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

@@

This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. The model contains \(m\)
  independent circular outcomes and \(w\) independent linear outcomes. The
  circular outcomes \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
  distribution. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,
  \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
  distribution [@sahu2003new]. Thus also in this model the linear outcome is
  defined constrained to be positive. Because the GPN distribution is modelled
  using a so-called augmented representation (see also the description of the
  CL-PN and CL-GPN models) it is convenient to use a similar tactic for
  modelling the multivariate skew normal distribution. As in
@mastrantonio2018joint dependence between the linear and circular outcome is
  created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
  and \(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
  distribution.\newline
\indent This means that we have a shared mean vector and variance-covariance
matrix for the linear and circular outcome(s), much like having multiple
outcomes in a MANOVA (multivariate analysis of variance) model. In our
regression extension of the GPN-SSN model we have \(i = 1, \dots, n\)
observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
covariates. The mean vector then becomes \(\boldsymbol{M}_i =
\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
\times (2m + w)\) matrix with regression coefficients and intercepts and
$\boldsymbol{x}_i$ is a $g + 1$ dimensional vector containing the value 1 to
estimate an intercept and the $g$ covariate values. This means that in contrast
to the other three models, we have to use the same set of predictors for the
circular and linear outcome.\newline
\indent For the teacher data, \(\boldsymbol{B} = \begin{bmatrix}
\beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} &
\beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\). The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integretion or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
linear outcomes in the GPN-SSN model are
\[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
\beta_{1_{s^{II}}}\text{SE}_i,\beta_{0_{s^{I}}} +
\beta_{1_{s^{I}}}\text{SE}_i),\]

@@

\noindent The GPN-SSN model thus allows us to assess the average type and
strength of interpersonal behavior through the parameters $\beta_{0_{s^{I}}}$,
$\beta_{0_{s^{II}}}$ and $\beta_{0_{y}}$ as well as the effect of self-efficacy
on type and strength of teacher behavior through the parameters,
$\beta_{1_{s^{I}}}$, $\beta_{1_{s^{II}}}$ and $\beta_{1_{y}}$. In addition,
because the type and strength of interpersonal behavior are modelled together
using a multivariate normal distribution we can through its variance-covariance
matrix also assess the dependence between the type and strength of interpersonal
behavior. \newline
\indent We estimate the model using MCMC methods. A detailed description of these
methods is given in the Supplementary Material.

@@

In the Supplementary Material we have described the starting values for the MCMC
procedures for the CL-PN, CL-GPN and GPN-SSN models, hence it remains to specify
the starting  values for the maximum likelihood based Abe-Ley model: \(\eta_0 =
0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda
= 0\). The initial number of iterations for the three MCMC samplers was set to
2000. After convergence checks via traceplots we concluded that some of the
parameters of the GPN-SSN model did not converge. Therefore we set the number of
iterations of the MCMC models to 20,000 and subtracted a burn-in of 5000 to
reach convergence (the Geweke diagnostics show absolute z-scores over 1.96 in
6\% of the estimated parameters). Note that we choose the same number of
iterations for all three models models estimated using MCMC prodedures to make
their comparison via the PLSL as fair as possible. Lastly, the predictor
\verb|SE| was centered before inclusion in the analysis as this allows the
intercepts to bear the classical meaning of average behavior.\newline
\indent Tables \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} show
the results for the four cylindrical models that were fit to the teacher data.
For the models estimated using MCMC methods, CL-PN, CL-GPN and GPN-SSN we show
descriptives of the posterior of the estimated parameters (posterior mode and
lower and upper bound of the 95\% highest posterior density (HPD) interval). For
the Abe-Ley model we show the maximum likelihood estimates of the parameters. To
compare the results of the four models we focus on the following aspects: the
estimated average scores (intercept) on the type and strength of interpersonal
behavior (1), the effect of self-efficacy on the type and strength of
interpersonal behavior (2), the dependence between the type and strength of
interpersonal behavior (3) and the model fit (4).

@@

The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley model and $\beta_{1_y}$

@@

conclusions on. The model CL-GPN and CL-PN fit the linear outcome best according

@@

Abe-Ley and GPN-SSN model respectively. For the CL-PN and Abe-Ley model we have

@@

the description of the CL-PN and CL-GPN models). For the CL-PN model the
inflection point is indicated with a square in Figure \ref{regline}. The
inflection point for the Abe-Ley model falls outside the bounds of the plot and

>>>>

In social sciences the use of cylindrical data is very common. Such data consist
of a linear and a circular component. @gurtman2011reasoning refers to such data
as vectors, with a directional measure (i.e., the circular component) and a
measure indicating the magnitude (i.e., the linear component). Many established
models in psychology are often referred to as circular or circumplex models, but
those models are cylindrical. Examples of such cylindrical models are the
interpersonal circle/circumplex
[@leary1957;@wiggins1996history;@wubbels2006interpersonal], the circumplex of
affect [@russell1980circumplex], the circumplex of human emotion
[@plutchik1997general] or the model of human values [@schwartz1992values].
\newline
\indent Also, many of the more recent types of data that are studied in psychology are
cylindrical. For example, research on human navigation uses data where distance
(i.e., linear component) and direction (i.e., the circular component) are of
interest [@chrastil2017rotational] or in eye-tracking, the saccade data also
consist of both the direction (i.e., the circular variable) and the duration
(i.e., the linear variable) (e.g., @rayner200935th). Apart from the social
sciences, data with a circular and linear outcome more commonly occur in
meteorology [@garcia2013exploring], ecology [@garcia2014test] or
marine research [@lagona2015hidden]
\newline
\indent Up until now researchers studying cylindrical data had to rely on linear
statistical methods to analyze their research results. However, lately more and more
of these researchers acknowledge that linear methods are not sufficient and call
for new methods [@gurtman2011reasoning; @pennings2017complexity;
@wright2009integrating] that take into account both the circular and the linear
component of these data.
\newline
\indent The aim in the present paper is twofold. Firstly, we intend to fill
the above mentioned gap in the literature by showing that the use of cylindrical
models can benefit the analysis of circumplex data and cylindrical data in
psychology in general. More specifically we will show these benefits for
interpersonal teacher data from the field of educational psychology. Apart from
modelling the dependence between the linear and circular component of a
cylindrical variable we would also like to predict the two components from a set
of covariates in a regression model. Our second aim therefore is to adapt
several existing cylindrical models in such a way that they include a regression
of both the linear and circular component of a cylindrical variable onto a set
of covariates. From now on we will therefore refer to the components of the
cylindrical variable as outcome components. These adapted cylindrical models are
then used  to analyse the teacher data.

@@

Data that consist of a linear variable and a circular variable is called
cylindrical data. A circular variable is different from the linear variable in
the sense that it is measured on a different scale. Figure \ref{circline} shows
the difference between a circular scale (right) and a linear scale (left). The
most important difference is that on a circular scale the datapoints 0\(^\circ\)
and 360\(^\circ\) are connected and in fact represent the same number while on a
linear scale the two ends, \(-\infty\) and \(\infty\), are not connected and
consequently the values 0\(^\circ\) and 360\(^\circ\) are located on different
places on the scale. Both circular data and cylindrical data require special
analysis methods due to this periodicity in the scale of a circular variable
(see e.g. @fisher1995statistical for an introduction to circular data and
@mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
elaborate overview).\newline
\indent In the literature, several methods have been put forward to model the
relation between the linear and circular component of a cylindrical variable.
Some of these are based on regressing the linear component onto the circular
component using the following type of relation: \[y = \beta_0 +
\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
linear component and \(\theta\) the circular component [@mardia1978model;
@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a
different way, e.g. by specifying a multivariate model for several linear and
circular variables and modelling their covariance matrix
[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For
example, @abe2017tractable introduce a cylindrical distribution based on a
Weibull distribution for the linear component and a sine-skewed von Mises
distribution for the circular component and link these through their respective
shape and concentration parameters. However, none of the methods that have been
proposed thus far include additional covariates onto which both the circular and
linear component are regressed.

@@

two-argument arctangent function in \eqref{PredVal}, where \(A\) represents a score on the Agency dimension and
\(C\) represents a score on the Communion dimension \footnote{The selection of

@@

bivariate space.}. Note that when placing a

@@

The teacher data was collected between 2010 and 2015 and contains several
repeated measures on the IPC of 161 teachers. Measurements were obtained using
the QTI and taken in different years and classes. For this paper we only
consider one measurement, the first occasion (2010) and largest class if data
for multiple classes were available. This results in a sample of 151 teachers.
In addition to the type of interpersonal behavior (IPC), the circular outcome,
and the strength of interpersonal behavior (IPC strength), the linear outcome, a
teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
covariate in the analysis. In previous research, in psychology and education it
has been shown that higher self-efficacy is related to the quality of
interpersonal interactions [@locke2007selfefficacy; @want2018selfefficacy].
After listwise deletion of missings ($3$ in total, only for the self-efficacy)
we have a sample of 148 teachers. Table \ref{Tableteacherdescriptives} shows
descriptives for the dataset. For the circular variable IPC we show sample
estimates for the circular mean $\bar{\theta}$ and concentration $\hat{\rho}$.
The circular concentration lies between 0, meaning the data is not concentrated
at all \emph{i.e.} spread over the entire circle, and 1, meaning all data is
concentrated at a single point on the circle. The population values of these
parameters are usually, and also in this paper, referred to as $\mu$ (circular
location) and $\kappa$ (circular concentration). For the linear variables
(strength IPC and SE) we show sample estimates of the linear mean and standard
deviation (sd). Figure \ref{dataplot} is a scatterplot showing the relation between
the linear and circular outcome of the teacher data for teachers with low SE
(below 1 sd below the mean), average SE (between 1 sd below and 1 sd above the
mean) and high SE (above 1 sd above the mean).

@@

One of the goals of this paper is to show the benefits of cylindrical methods
for the analysis of circumplex data and cylindrical data in psychology in
general. To do so we decide to focus on four cylindrical models. The models were
selected for their relatively low complexity and the ease with which a
regression structure could be incorporated. But also because they show different
ways of modelling the linear and circular outcome and thereby illustrate a
wider range of cylindrical models available in the literature. The
cylindrical models contain a set of $q$ predictors $\boldsymbol{x} = x_1, \dots,
x_q$ and $p$ predictors $\boldsymbol{z} = z_1, \dots, z_p$ for the linear and
circular outcomes, \(Y\) and \(\Theta\), respectively. The first two models are
based on a construction by @mastrantonio2015bayesian, while the other models are
extensions of the models from @abe2017tractable and @mastrantonio2018joint. The
four cylindrical  models are introduced separately in the subsections below.
However, to provide a more succinct overview and comparison of the four models, Table
\ref{TableModels} gives an overview of the similarities and differences between
the models.

\begin{sidewaystable}[h]
\centering
\caption{Comparison of the four cylindrical regression models} 
\begin{tabular}{lllll}
  \noalign{\smallskip}\hline\noalign{\smallskip}
\multicolumn{1}{l}{Aspect} & CL-PN & CL-GPN  & Abe-Ley  & GPN-SSN \\ \hline\noalign{\smallskip}
$\Theta$ & &&&\\
$\:\:$Distribution& PN & GPN & Sine-skewed vM & GPN\\
$\:\:$Domain & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$ & $[0, 2\pi)$\\
$\:\:$Shape & symmetric, & asymmetric, & asymmetric, & asymmetric, \\
            & unimodal  & multimodal & unimodal   & multimodal \\\hline\noalign{\smallskip}
$Y$& &&&\\
$\:\:$Distribution & Normal & Normal & Weibull & skewed-Normal\\
$\:\:$Domain & $(-\infty, + \infty)$ & $(-\infty, + \infty)$ & $(0, + \infty)$ & $(-\infty, + \infty)$\\
$\:\:$Shape & symmetric, & symmetric, & asymmetric, & asymmetric, \\
            & unimodal  & unimodal  & unimodal   & unimodal\\\hline\noalign{\smallskip}
$\Theta$-$Y$ dependence &                                   &                                   & & \\
                        & $y$ regressed on                  & $y$ regressed on                  & $\alpha$ and $\kappa$ & multivariate \\
                        & $\sin(\theta)$ and $\cos(\theta)$ & $\sin(\theta)$ and $\cos(\theta)$ & & distribution\\\hline
\multicolumn{5}{l}{Note: PN and GPN refer to the projected normal and general projected normal distribution.}\\
\multicolumn{5}{l}{vM refers to the von-Mises distribution}\\

@@

circular outcome. It is a variable that in contrast to \(\Theta\) is not

@@

\indent For both the PN and GPN distributions the circular location parameter

@@

integration or Monte Carlo methods because a closed form expression for the mean

@@

\ref{circregline}). By default, the parameters from the PN and GPN models do not

@@

& \xi\\ \xi & 1 \end{bmatrix}\) where \(\xi,\tau \in (-\infty, +\infty)\), allowing for multimodality and
asymetry/skewness.\newline

@@

the scores on the Agency and Communion dimensions respectively.\newline

@@

\begin{figure}[]
  \includegraphics[width = \textwidth]{Plots/circregline.pdf}
  \caption{Circular regression line for the relation between a covariate and a circular outcome with the data the regression line was fit to. The square indicates the inflection point of the regression line.}
  \label{circregline}
\end{figure}

@@

This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. It concerns a combination
of a Weibull distribution, with scale parameter \(\nu>0\) and shape parameter
\(\alpha\), for the linear outcome and a sine-skewed von Mises distribution,
with location parameter \(\mu\in [0, 2\pi)\), concentration parameter
\(\kappa>0\) and skewness \(\lambda \in [-1,1]\), for the circular outcome. In
contrast to the CL-PN and CL-GPN models, the linear outcome \(Y\) is in this
model defined only on the positive real half-line \([0, + \infty)\) and thus can
not be negative.\newline
\indent In this model we predict the linear scale parameter and circular
location parameter, both of which we can express in terms of
covariates: \(\hat{\nu}_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\hat{\mu}_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\gamma_0\) is the intercept. The parameter \(\beta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\).\newline
\indent For the teacher data, the predicted values for the circular outcome in
the Abe-Ley model are: \[\hat{\theta}_{i} = \hat{\mu}_i = \beta_0 + 2 *
\tan^{-1}(\beta_1\text{SE}_i).\] We do not directly predict the linear outcome.
The conditional distribution for the linear outcome is Weibull, meaning that we
can use methods from survival analsis to interpret the effect of a predictor. In
survival analysis a 'survival' function is used in which time is plotted against
the probability of survival of subjects suffering from a specific medical
condition. In the teacher data we can thus compute the probability of a teacher
having a specific strength on the IPC. This probability is computed using the
'survival-function' defined as  \[\exp(-\alpha
y_i^{\hat{\nu}_i(1-\tanh(\kappa)\cos(\theta_i - \hat{\mu}_i))^{1/\alpha}}),\]
with \(\hat{\nu}_i = \exp(\gamma_0 + \gamma_1\mbox{SE}_i)\). From the survival
function we also see that the circular concentration parameter \(\kappa\) and
the linear shape parameter \(\alpha\) regulate the circular-linear dependence in
the Abe-Ley model. The Abe-Ley model thus allows us to assess the average type
and strength of interpersonal behavior through the parameters $\beta_{0}$ and
$\gamma_0$ as well as the effect of self-efficacy on type and strength of
teacher behavior through the parameters $\beta_{1}$ and $\gamma_1$.\newline
\indent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

@@

This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. The model contains \(m\)
  independent circular outcomes and \(w\) independent linear outcomes. The
  circular outcomes \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
  distribution. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots,
  \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
  distribution [@sahu2003new]. Because the GPN distribution is modelled
  using a so-called augmented representation (see also the description of the
  CL-PN and CL-GPN models) it is convenient to use a similar tactic for
  modelling the multivariate skew normal distribution. As in
@mastrantonio2018joint, dependence between the linear and circular outcome is
  created by modelling the augmented representations of \(\boldsymbol{\Theta}\)
  and \(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
  distribution.\newline
\indent This means that we have a shared mean vector and variance-covariance
matrix for the linear and circular outcome(s), much like having multiple
outcomes in a MANOVA (multivariate analysis of variance) model. In our
regression extension of the GPN-SSN model we have \(i = 1, \dots, n\)
observations of \(m\) circular outcomes, \(w\) linear outcomes and \(g\)
covariates. The mean vector then becomes \(\boldsymbol{M}_i =
\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
\times (2m + w)\) matrix with regression coefficients and intercepts and
$\boldsymbol{x}_i$ is a $g + 1$ dimensional vector containing the value 1 to
estimate an intercept and the $g$ covariate values. This means that in contrast
to the other three models, we have to use the same set of predictors for the
circular and linear outcome.\newline
\indent For the teacher data, \(\boldsymbol{B} = \begin{bmatrix}
\beta_{0_{s^{I}}} & \beta_{0_{s^{II}}} & \beta_{0_{y}}\\ \beta_{1_{s^{I}}} &
\beta_{1_{s^{II}}} & \beta_{1_{y}} \end{bmatrix}\). The predicted circular\footnote{Note that for the  GPN-SSN model the predicted circular outcome also depends on the variance-covariance matrix and the circular predicted values should be computed using numerical integration or Monte Carlo methods because a closed form expression for the mean direction is not available.} and
linear outcomes in the GPN-SSN model are
\[\hat{\theta_i} = \mbox{atan2}(\beta_{0_{s^{II}}} +
\beta_{1_{s^{II}}}\text{SE}_i,\beta_{0_{s^{I}}} +
\beta_{1_{s^{I}}}\text{SE}_i),\]

@@

\noindent The GPN-SSN model thus allows us to assess the average type and
strength of interpersonal behavior through the parameters $\beta_{0_{s^{I}}}$,
$\beta_{0_{s^{II}}}$ and $\beta_{0_{y}}$ as well as the effect of self-efficacy
on type and strength of teacher behavior through the parameters
$\beta_{1_{s^{I}}}$, $\beta_{1_{s^{II}}}$ and $\beta_{1_{y}}$. In addition,
because the type and strength of interpersonal behavior are modelled together
using a multivariate normal distribution we can, through its variance-covariance
matrix, also assess the dependence between the type and strength of interpersonal
behavior. \newline
\indent We estimate the model using MCMC methods. A detailed description of these
methods is given in the Supplementary Material.

@@

In the Supplementary Material we have described the starting values for the MCMC
procedures for the CL-PN, CL-GPN and GPN-SSN models, hence it remains to specify
the starting  values for the maximum likelihood based Abe-Ley model: \(\eta_0 =
0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda
= 0\). The initial number of iterations for the three MCMC samplers was set to
2000. After convergence checks via traceplots we concluded that some of the
parameters of the GPN-SSN model did not converge. Therefore we set the number of
iterations of the MCMC models to 20,000 and subtracted a burn-in of 5000 to
reach convergence (the Geweke diagnostics show absolute z-scores over 1.96 in
6\% of the estimated parameters). Note that we choose the same number of
iterations for all three models estimated using MCMC procedures to make
their comparison via the PLSL as fair as possible. Lastly, the predictor
\verb|SE| was centered before inclusion in the analysis as this allows the
intercepts to bear the classical meaning of average behavior.\newline
\indent Tables \ref{tab:estCLGPN}, \ref{tab:estAL} and \ref{tab:estCLGPNM} show
the results for the four cylindrical models that were fit to the teacher data.
For the models estimated using MCMC methods, CL-PN, CL-GPN and GPN-SSN, we show
descriptives of the posterior of the estimated parameters (posterior mode and
lower and upper bound of the 95\% highest posterior density (HPD) interval). For
the Abe-Ley model we show the maximum likelihood estimates of the parameters. To
compare the results of the four models we focus on the following aspects: the
estimated average scores (intercept) on the type and strength of interpersonal
behavior (1), the effect of self-efficacy on the type and strength of
interpersonal behavior (2), the dependence between the type and strength of
interpersonal behavior (3) and the model fit (4).

@@

The parameters $\gamma_1$ in the CL-PN, CL-GPN, Abe-Ley models and $\beta_{1_y}$

@@

conclusions on. The models CL-GPN and CL-PN fit the linear outcome best according

@@

Abe-Ley and GPN-SSN model respectively. For the CL-PN and Abe-Ley models we have

@@

the description of the CL-PN and CL-GPN models for a detailed explanation of
circular regression lines). For the CL-PN model the inflection point is
indicated with a square in Figure \ref{regline}. The inflection point for the
Abe-Ley model falls outside the bounds of the plot and is therefore not
